MG7: Configurable and scalable 16S metagenomics data analysis -- new methods optimized for massive cloud computing

### 2015-09-14 docs/paper/out/paper.tex

\href{https://github.com/ohnosequences/statika}{Statika} is a Scala
library developed by the first and last authors which serves as a way of
defining and composing machine behaviors statically. The main component
are \textbf{bundles}. Each bundle declares a sequence of computations
(its behavior) which will be executed in an \textbf{environment}. A
bundle can \emph{depend} on other bundles, and when being executed by an
environment, its DAG of dependencies is linearized and run in sequence.
In our use, bundles correspond to what an EC2 instance should do and an
environment to an image (AMI: \textbf{A}mazon \textbf{M}achine
\textbf{I}mage) which prepares the basic configuration, downloads the
Scala code and runs it.

@@

\href{https://github.com/ohnosequences/loquat}{Loquat} is a library
developed by the first, second and last authors designed for the
execution of embarrassingly parallel tasks using S3, SQS and EC2.

A \textbf{loquat} executes a process with explicit input and output
datasets (declared using the \emph{Datasets} library described above).
Workers (EC2 instances) read from an SQS queue the S3 locations for both
input and output data; then they download the input to local files, and
pass these file locations to the process to be executed. The output is
then put in the corresponding S3 locations.

@@

Both worker and manager instances are Statika bundles. In the case of
the worker, it can declare any dependencies needed to perform its task:
other tools, libraries, or data.

@@

Generic type-safe parsers returning an heterogeneous record of BLAST
output fields are also available, together with output data defined
using \emph{Datasets} which have a reference to the exact BLAST command
options which yielded that output. This let us provide generic parsers
for BLAST output which are guaranteed to be correct, for example.

In the same spirit as for BLAST, we implemented a type-safe EDSL for
FLASH expressions and their execution, sporting features equivalent to
those outlined for the BLAST EDSL.

@@

We use EC2, S3 and SQS through a Scala wrapper of the official
\href{https://aws.amazon.com/sdk-for-java/}{AWS Java SDK},
\href{https://github.com/ohnosequences/aws-scala-tools/???}{ohnosequences/aws-scala-tools
0.13.2}. This uses version \texttt{1.10.9} of the AWS Java SDK.

@@

MG7 uses
\href{https://github.com/statika/statika/releases/tag/v0.2.0}{ohnosequences/statika
2.0.0} for specifying the configuration and behavior of EC2 instances.

@@

MG7 uses
\href{https://github.com/ohnosequences/datasets/releases/tag/v0.2.0}{ohnosequences/datasets
0.2.0} for specifying input and output data, their type and their
location.

@@

MG7 uses
\href{https://github.com/ohnosequences/loquat/releases/tag/v2.0.0}{ohnosequences/loquat
2.0.0} for the specification of data processing tasks and their
execution using AWS resources.

@@

MG7 uses
\href{https://github.com/ohnosequences/blast/releases/tag/v0.2.0}{ohnosequences/blast
0.2.0}. The BLAST version used is \texttt{2.2.31+}

@@

MG7 uses
\href{https://github.com/ohnosequences/flash/releases/tag/v0.1.0}{ohnosequences/flash
0.1.0}. The FLASH version used is \texttt{?.?.?}

>>>>

\href{https://github.com/ohnosequences/statika}{Statika} is a Scala
library developed by the first and last authors which serves as a way of
defining and composing machine behaviors statically. The main component
are \textbf{bundles}. Each bundle declares a sequence of computations
(its behavior) which will be executed in an \textbf{environment}. A
bundle can \emph{depend} on other bundles, and when being executed by an
environment, its DAG (Directed Acyclic Graph) of dependencies is
linearized and run in sequence. In our use, bundles correspond to what
an EC2 instance should do and an environment to an AMI (Amazon Machine
Image) which prepares the basic configuration, downloads the Scala code
and runs it.

@@

\href{https://github.com/ohnosequences/loquat}{Loquat} is a library
developed by the first, second and last authors designed for the
execution of embarrassingly parallel tasks using S3, SQS and EC2 Amazon
services.

A \emph{loquat} executes a process with explicit input and output
datasets (declared using the \emph{Datasets} library described above).
Workers (EC2 instances) read from an SQS queue the S3 locations for both
input and output data; then they download the input to local files, and
pass these file locations to the process to be executed. The output is
then put in the corresponding S3 locations.

@@

Both worker and manager instances are \emph{Statika} bundles. In the
case of the worker, it can declare any dependencies needed to perform
its task: other tools, libraries, or data.

@@

Generic type-safe parsers returning a heterogeneous record of BLAST
output fields are also available, together with output data defined
using \emph{Datasets} which have a reference to the exact BLAST command
options which yielded that output. This let us provide generic parsers
for BLAST output which are guaranteed to be correct.

In the same spirit as for BLAST, we implemented a type-safe eDSL for
FLASH expressions and their execution, supporting features equivalent to
those outlined for the BLAST eDSL.

@@

We use EC2, S3 and SQS through a Scala wrapper of the official
\href{https://aws.amazon.com/sdk-for-java/}{AWS Java SDK},
\href{https://github.com/ohnosequences/aws-scala-tools/releases/tag/v0.13.2}{ohnosequences/aws-scala-tools
\texttt{0.13.2}}. This uses version \texttt{1.9.25} of the AWS Java SDK.

@@

MG7 uses
\href{https://github.com/statika/statika/releases/tag/v0.2.0}{ohnosequences/statika
\texttt{2.0.0}} for specifying the configuration and behavior of EC2
instances.

@@

MG7 uses
\href{https://github.com/ohnosequences/datasets/releases/tag/v0.2.0}{ohnosequences/datasets
\texttt{0.2.0}} for specifying input and output data, their type and
their location.

@@

MG7 uses
\href{https://github.com/ohnosequences/loquat/releases/tag/v2.0.0}{ohnosequences/loquat
\texttt{2.0.0}} for the specification of data processing tasks and their
execution using AWS resources.

@@

MG7 uses
\href{https://github.com/ohnosequences/blast/releases/tag/v0.2.0}{ohnosequences/blast
\texttt{0.2.0}}. The BLAST version used is \texttt{2.2.31+}

@@

MG7 uses
\href{https://github.com/ohnosequences/flash/releases/tag/v0.1.0}{ohnosequences/flash
\texttt{0.1.0}}. The FLASH version used is \texttt{1.2.11}
